\documentclass{article}

% Language setting
% Replace `english' with e.g. `spanish' to change the document language
\usepackage[english]{babel}

% Set page size and margins
% Replace `letterpaper' with `a4paper' for UK/EU standard size
\usepackage[letterpaper,top=2cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage[colorlinks=true, allcolors=blue]{hyperref}

\title{PS1 - ECON 5253}
\author{Hannah Bermudez}

\begin{document}
\maketitle

\section{Introduction}

\hspace*{2em} I am mainly interested in how psychological and behavioral elements influence decision making and I think it shares considerable overlap with economics. But I was initially drawn to this class because I want to learn more about data analytics, machine learning, and how to use statistical techniques to streamline my ability to code and analyze data sets effectively. Additionally, I do not have a very comprehensive foundation in terms of data manipulation or coding, so I am also trying to bolster my overall understanding through hands-on experience and the pressure of a classroom. 

In terms of the project for this class, I am currently working on my first-year summer paper for my program and am about to run an experiment. The experiment will focus on the effect of priming and professional skepticism on sequential belief revision. The original paper on which I am basing my experiment did not focus so much on the generalization of their experiments, but on the internal validity of their findings. I want to know whether I get similar results in a more realistic audit environment. Therefore, it will be two parts: a replication and an expansion. I think it would be interesting to try and leverage the tools used in this class to analyze the data I will compile. 

My goals for this class are to learn as much as I can and use what I have learned to enhance my ability to analyze data sets and apply the techniques to my future research endeavors. I will not graduate for a while, but I want to eventually do well in academia and teach students. I think this class is a good start towards that goal.

\section{Equation}

\[ a^2 + b^2 = c^2 \]

\end{document}